---
title: "homework2"
author: Na Yun Cho
output: pdf_document
---

```{r}
library(caret)
library(splines)
library(mgcv)
library(pdp)
library(earth)
library(tidyverse)
library(ggplot2)
```



## (a) Exploratory data analysis
```{r}
college = read.csv("./data/data.csv")
college1 <- college[-125,]

college2 = data.matrix(college1, rownames.force = NA)
x <- college2 [ ,-c(1,9)]
y <- college2 [ , 9]


theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5)
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
featurePlot(x, y, plot = "scatter", labels = c("","Y"),
            type = c("p"), layout = c(4, 2))
```

Interpretation: From this exploratory data analysis, I could see that the predictors 'F.Undergrad', 'P.Undergrad', 'Apps', 'Accept', 'Enroll', 'Terminal', and 'Books' show a relatively non-linear trend compared to other predictors. The predictors 'Top25perc', 'Room.Board', 'Top10perc', 'perc.alumni', 'Grad.Rate', 'Expend', and 'PhD' showed a generally increasing trend that looks quite linear. On the other hand, 'S.F.Ratio' and 'Personal' seemed to show a slightly decreasing trend that is quite linear. To check the associations of each predictor with the outcome 'Outstate' in more detail, further analyses would have to be done. 



## (b) Fit a smoothing spline model using 'Terminal' as the only predictor
```{r}
# using GCV method
fit.ss <- smooth.spline(college1$Terminal, college1$Outstate)
fit.ss$df

Terminallims <- range(college1$Terminal)
Terminal.grid <- seq(from = Terminallims[1],to = Terminallims[2])

pred.ss <- predict(fit.ss, x = Terminal.grid)
pred.ss.df <- data.frame(pred = pred.ss$y, Terminal = Terminal.grid)

p <- ggplot(data= college1, aes(x = Terminal, y = Outstate)) +geom_point(color = rgb(0.2, 0.4, 0.2, 0.5))
p + geom_line(aes(x = Terminal, y =pred), data = pred.ss.df, color = rgb(0.8, 0.1, 0.1, 1)) +theme_bw()


#Using LOOCV method
fit.ss <- smooth.spline(college1$Terminal, college1$Outstate, cv = TRUE)
fit.ss$df

Terminallims <- range(college1$Terminal)
Terminal.grid <- seq(from = Terminallims[1],to = Terminallims[2])

pred.ss <- predict(fit.ss, x = Terminal.grid)
pred.ss.df <- data.frame(pred = pred.ss$y, Terminal = Terminal.grid)

p <- ggplot(data= college1, aes(x = Terminal, y = Outstate)) +geom_point(color = rgb(0.2, 0.4, 0.2, 0.5))
p + geom_line(aes(x = Terminal, y =pred), data = pred.ss.df, color = rgb(0.8, 0.1, 0.1, 1)) +theme_bw()


#Using arbitrary lambda values
#Using lambda = 10
fit.ss <- smooth.spline(college1$Terminal, college1$Outstate, lambda=10)
fit.ss$df

Terminallims <- range(college1$Terminal)
Terminal.grid <- seq(from = Terminallims[1],to = Terminallims[2])

pred.ss <- predict(fit.ss, x = Terminal.grid)
pred.ss.df <- data.frame(pred = pred.ss$y, Terminal = Terminal.grid)

p <- ggplot(data= college1, aes(x = Terminal, y = Outstate)) +geom_point(color = rgb(0.2, 0.4, 0.2, 0.5))
p + geom_line(aes(x = Terminal, y =pred), data = pred.ss.df, color = rgb(0.8, 0.1, 0.1, 1)) +theme_bw()

#Using lambda = 0.5
fit.ss <- smooth.spline(college1$Terminal, college1$Outstate, lambda=0.5)
fit.ss$df

Terminallims <- range(college1$Terminal)
Terminal.grid <- seq(from = Terminallims[1],to = Terminallims[2])

pred.ss <- predict(fit.ss, x = Terminal.grid)
pred.ss.df <- data.frame(pred = pred.ss$y, Terminal = Terminal.grid)

p <- ggplot(data= college1, aes(x = Terminal, y = Outstate)) +geom_point(color = rgb(0.2, 0.4, 0.2, 0.5))
p + geom_line(aes(x = Terminal, y =pred), data = pred.ss.df, color = rgb(0.8, 0.1, 0.1, 1)) +theme_bw()

#Using lambda = 0.001
fit.ss <- smooth.spline(college1$Terminal, college1$Outstate, lambda=0.001)
fit.ss$df

Terminallims <- range(college1$Terminal)
Terminal.grid <- seq(from = Terminallims[1],to = Terminallims[2])

pred.ss <- predict(fit.ss, x = Terminal.grid)
pred.ss.df <- data.frame(pred = pred.ss$y, Terminal = Terminal.grid)

p <- ggplot(data= college1, aes(x = Terminal, y = Outstate)) +geom_point(color = rgb(0.2, 0.4, 0.2, 0.5))
p + geom_line(aes(x = Terminal, y =pred), data = pred.ss.df, color = rgb(0.8, 0.1, 0.1, 1)) +theme_bw()


```

## (c) Fit a GAM model using all the predictors
```{r}
gam.m1 <- gam(Outstate~ Apps+Accept+Enroll+Top10perc+Top25perc+F.Undergrad+P.Undergrad+Room.Board+Books
              +Personal+Terminal+PhD+S.F.Ratio+perc.alumni+Expend+Grad.Rate, data = college1)

gam.m2 <- gam(Outstate~ Apps+Accept+Enroll+Top10perc+Top25perc+F.Undergrad+P.Undergrad+Room.Board+Books
              +Personal+s(Terminal)+PhD+S.F.Ratio+perc.alumni+Expend+Grad.Rate, data = college1)

gam.m3 <- gam(Outstate~ s(Apps)+s(Accept)+s(Enroll)+Top10perc+Top25perc+s(F.Undergrad)+s(P.Undergrad)
              +Room.Board+s(Books)+Personal+s(Terminal)+PhD+S.F.Ratio+perc.alumni+Expend+Grad.Rate,
              data = college1)

gam.m4 <- gam(Outstate~ s(Apps)+s(Accept)+ s(Enroll)+Top10perc+Top25perc+te(F.Undergrad,P.Undergrad)
              +Room.Board+s(Books)+Personal+s(Terminal)+PhD+S.F.Ratio+perc.alumni+Expend+Grad.Rate, 
              data = college1)

anova(gam.m1, gam.m2, gam.m3, gam.m4, test = "F")


plot(gam.m2)
plot(gam.m3)
vis.gam(gam.m4, view=c("F.Undergrad","P.Undergrad"), color = "topo")
vis.gam(gam.m4, view=c("F.Undergrad","P.Undergrad"), color = "topo", plot.type = "contour")
```

Description: I have fitted multiple GAM models using all the predictors. From running the ANOVA test and checking the p-values of the result output, I could see that the third GAM model(gam.m3) shows the best fit. 
A plot of the second GAM model(gam.m2) showed the plot of the s function of the 'Terminal' predictor. The plot was neither noticeably increasing/decreasing until around a 'Terminal' value of 80, and then showed an increasing trend as 'Terminal' value increased. The credible interval spread out towards the ends of the plot.
I have also plotted all the s functions of the third GAM model. 
The plot of the s function of 'Accept' showed a slightly increasing trend and the credible interval spread out quite a lot as the values of the predictor increased. 
However, the plot of the s function of 'Enroll' showed a slgitly decreasing trend and the credible interval spread out quite a lot as well as the values of the predictor increased. 
For all the other plots of the s functions of each of their corresponding predictors, the plots did not noticeably increase or decrease. 
I have also plotted the te function of 'F.Undergrad' and 'P.Undergrad' predictors, and the different colors indicate the magnitude of the functions. 


## (d)Train a MARS model using all the predictors 
```{r}
ctrl1 <- trainControl(method = "cv", number = 10)

mars_grid <-expand.grid(degree = 1:3,
                        nprune = 2:20)
set.seed(1)
mars.fit <- train(x, y,
                  method = "earth",
                  tuneGrid = mars_grid,
                  trControl = ctrl1)
ggplot(mars.fit)
mars.fit$bestTune
coef(mars.fit$finalModel) 
```
For the final model, the nprune value is 14 and degree is 1. 
The final model is Intercept -240.692034 * h(Grad.Rate-97) -24.5321278 * h(97- Grad.Rate) -0.3593198 * h(F.Undergrad-1355) -1.6238125 * h(1355-F.Undergrad). 
The hinge functions are h(Grad.Rate-97), h(97- Grad.Rate), and h(F.Undergrad-1355), h(1355-F.Undergrad). 

## (d)Present the partial dependence plots 
```{r}
p1<- pdp::partial(mars.fit, pred.var = c("Grad.Rate"), grid.resolution = 10) %>% autoplot()
p2 <- pdp::partial(mars.fit, pred.var = c("F.Undergrad"), grid.resolution = 10) %>% autoplot()

grid.arrange(p1, p2, ncol = 2)
```

